Search-Based Test Data Generation for JavaScript Functions that Interact with the DOM \thanks{Identify applicable funding agency here. If none, delete this.}

### 2018-05-11 doc/issre2018/framework.tex
Near final version of the issre 2018 paper

\documentclass[sigconf,review]{acmart}

\usepackage[inline]{enumitem}
\usepackage{tablefootnote}
\usepackage{syntax}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{booktabs} % For formal tables
\usepackage{tikz}
\usetikzlibrary{arrows,shadows,positioning,decorations.pathmorphing}
\usepackage{pgf-umlsd}
\usepackage{tikz-uml}
\usepackage{hyperref}
\usepackage{upquote}
\usepackage{listings}
\usepackage{lsthaskell}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{mdframed}
\usepackage{algorithmic}
\usepackage{longtable}
\usepackage{subcaption}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepackage[titlenotnumbered, linesnumbered, ruled]{algorithm2e} % language for algorithm description
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = red, %Colour of internal links
  citecolor    = green %Colour of citations
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\definecolor{ltblue}{rgb}{0,0.4,0.4}
\definecolor{dkblue}{rgb}{0,0.1,0.6}
\definecolor{dkgreen}{rgb}{0,0.4,0}
\definecolor{dkviolet}{rgb}{0.3,0,0.5}
\definecolor{dkred}{rgb}{0.5,0,0}

\definecolor{lightgray}{rgb}{0.95, 0.95, 0.95}
\definecolor{darkgray}{rgb}{0.4, 0.4, 0.4}
%\definecolor{purple}{rgb}{0.65, 0.12, 0.82}
\definecolor{editorGray}{rgb}{0.95, 0.95, 0.95}
\definecolor{editorOcher}{rgb}{1, 0.5, 0} % #FF7F00 -> rgb(239, 169, 0)
\definecolor{editorGreen}{rgb}{0, 0.5, 0} % #007C00 -> rgb(0, 124, 0)
\definecolor{orange}{rgb}{1,0.45,0.13}		
\definecolor{olive}{rgb}{0.17,0.59,0.20}
\definecolor{brown}{rgb}{0.69,0.31,0.31}
\definecolor{purple}{rgb}{0.38,0.18,0.81}
\definecolor{lightblue}{rgb}{0.1,0.57,0.7}
\definecolor{lightred}{rgb}{1,0.4,0.5}
% CSS
\lstdefinelanguage{CSS}{
  keywords={color,background-image:,margin,padding,font,weight,display,position,top,left,right,bottom,list,style,border,size,white,space,min,width, transition:, transform:, transition-property, transition-duration, transition-timing-function},	
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]",
  alsoletter={:},
  alsodigit={-}
}

% JavaScript
\lstdefinelanguage{JavaScript}{
  morekeywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, for},
  morecomment=[s]{/*}{*/},
  morecomment=[l]//,
  morestring=[b]",
  morestring=[b]'
}

\lstdefinelanguage{HTML5}{
  language=html,
  sensitive=true,	
  alsoletter={<>=-},	
  morecomment=[s]{<!-}{-->},
  morestring=[b]'
  tag=[s],
  otherkeywords={
  % General
  >,
  % Standard tags
	<!DOCTYPE,
  </html, <html, <head, <title, </title, <style, </style, <link, </head, <meta, />,
	% body
	</body, <body,
	% Divs
	</div, <div, </div>, 
	% Paragraphs
	</p, <p, </p>,
	% scripts
	</script, <script,
  % More tags...
  <canvas, /canvas>, <svg, <rect, <animateTransform, </rect>, </svg>, <video, <source, <iframe, </iframe>, </video>, <image, </image>, <header, </header, <article, </article
  },
  ndkeywords={
  % General
  =,
  % HTML attributes
  charset=, src=, id=, width=, height=, style=, type=, rel=, href=,
  % SVG attributes
  fill=, attributeName=, begin=, dur=, from=, to=, poster=, controls=, x=, y=, repeatCount=, xlink:href=,
  % properties
  margin:, padding:, background-image:, border:, top:, left:, position:, width:, height:, margin-top:, margin-bottom:, font-size:, line-height:,
	% CSS3 properties
  transform:, -moz-transform:, -webkit-transform:,
  animation:, -webkit-animation:,
  transition:,  transition-duration:, transition-property:, transition-timing-function:,
  }
}

\lstdefinelanguage{HTML5}{
  language=html,
  tagstyle=\color{blue},
  markfirstintag=true,
  sensitive=true,
  alsoletter={<>=-},	
  morecomment=[s]{<!--}{-->},
  % tag=[s],
  %% morestring=[b]'',
  morestring=[b]',
  ndkeywords={
    % General
    =,
    % HTML attributes
    id=, class= 
  }
}

\lstdefinestyle{htmlcssjs} {%
  % General design
  backgroundcolor=\color{editorGray},
  basicstyle={\scriptsize\ttfamily},   
  frame=tb,
  % line-numbers
  xleftmargin={0.75cm},
  numbers=left,
  stepnumber=1,
  firstnumber=1,
  numberfirstline=true,	
  % Code design
  identifierstyle=\color{black},
  keywordstyle=\color{blue}\bfseries,
  ndkeywordstyle=\color{editorGreen}\bfseries,
  stringstyle=\color{editorOcher}\ttfamily,
  commentstyle=\color{brown}\ttfamily,
  % Code
  language=JavaScript,
  alsolanguage=HTML5,
  alsodigit={.:;},	
  tabsize=2,
  showtabs=false,
  showspaces=false,
  showstringspaces=false,
  extendedchars=true,
  breaklines=true,
  escapechar=|c
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\listfiles % used to output the versions of all files being used by the late

\newcommand{\Server}{\emph{Server}\xspace}
\newcommand{\Client}{\emph{Client}\xspace}
\newcommand{\User}{\emph{User}\xspace}
\newcommand{\Validator}{\emph{Validator}\xspace}
\newcommand{\thenBr}{\text{then}}
\newcommand{\elseBr}{\text{else}}
\newcommand{\inFor}{\text{inFor}}
\newcommand{\inWhile}{\text{while}}
\newcommand{\un}[1]{\underline{#1}}
\newcommand{\Random}{\mathbb{R}}
\newcommand{\Genetic}{\mathbb{G}}
\newcommand{\RGenetic}{\mathbb{G}_{10}}
\newcommand{\Confix}{\emph{Confix}}
\newcommand{\Jalangi}{\emph{Jalangi}}
\newcommand{\Jedi}{\emph{JEDI}}

% Copyright
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\acmDOI{10.475/123_4}

% ISBN
%\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[ISSTA 2018]{ACM SIGSOFT International Symposium on Software Testing and Analysis}{16-22 July, 2018}{Amsterdam, The Netherlands}
%% \acmYear{1997}                  
%% \copyrightyear{2016}

%% \acmPrice{15.00}

\begin{document}
\title[Search-Based Test Data Generation for JavaScript Functions...]{Search-Based Test Data Generation for JavaScript Functions that Interact with the DOM}

\author{Alexander Elyasov}
\affiliation{%
  \institution{Utrecht University, The Netherlands} 
}
\email{a.elyasov@uu.nl}

\author{Wishnu Prasetya}
\affiliation{%
  \institution{Utrecht University, The Netherlands}
}
\email{s.w.b.prasetya@uu.nl}

\author{Jurriaan Hage}
\affiliation{%
  \institution{Utrecht University, The Netherlands}
}
\email{j.hage@uu.nl}

\begin{abstract}
The popularity of JavaScript is continuously growing. Together with HTML and CSS, it forms the core technology stack for web development. Because of the dynamic nature and complex interplay with HTML, JS applications are often error-prone and vulnerable. Despite the active research efforts to devise intricate static and dynamic analyses to facilitate JS testing, the problem of test data generation for JS applications interacting with the DOM has not yet been addressed. In this paper, we present a novel \textbf{J}avascript \textbf{E}volution-based testing framework with \textbf{D}OM as an \textbf{I}nput, called \emph{JEDI}. In order to reach a target branch, it applies genetic search for relevant input parameters of the JS function including the global DOM state. We conducted an empirical evaluation to study the effectiveness and efficiency of our testing framework. It shows that the \emph{genetic with restart} algorithm, proposed in this paper, is able to achieve complete branch coverage for all experimental subjects, taking on average 19 seconds per branch.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011074.10011099.10011102.10011103</concept_id>
<concept_desc>Software and its engineering~Software testing and debugging</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011074.10011784</concept_id>
<concept_desc>Software and its engineering~Search-based software engineering</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[300]{Software and its engineering~Software testing and debugging}
\ccsdesc[300]{Software and its engineering~Search-based software engineering}

\keywords{JavaScript, search-based testing, branch coverage}

\maketitle

%% ------------------------------------------------------------------------
\section{Introduction}
\label{sec.intro}
%% ------------------------------------------------------------------------

According to the \emph{StackOverflow} Developer Survey 2018 \cite{stackoverflow2018}, JavaScript has been recognized for the sixth time in row as the most commonly used programming language. It is also the most popular language on \emph{GitHub}~\cite{guthub2017} based on the number of opened pull requests (2.3 million). Such an ever increasing JS success can be explained by its native browsers support, and the rise in popularity of the JS run-time environment \emph{Node.js}. On the other hand, due to of implicit type conversion, lack of static types, and extensive support of dynamic language constructs~\cite{richards2010analysis}, JS applications are often error-prone~\cite{frolin:TSE16}, hard to analyze~\cite{andreasen2017survey, sun2017analysis} and test~\cite{mesbah2015advances}. Thus, it is not surprising that once in a while critical bugs manage to break through~\cite{bugstories2017} into the web applications by Facebook, Twitter etc.

Systematic unit testing is basic and, at the same time, the most effective technique for improving software quality. In practice, there are numerous of JS frameworks --- Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma, QUnit etc. --- intended to assist developers with the automation of various unit testing activities such as design, execution and monitoring. Despite the apparent abundance of the JS testing tools, the developers' satisfaction rate is still relative low~\cite{stateJS2017} (3.2 out of 5), suggesting the need for improvements. One of such areas is \emph{test data generation}. It pursues the goal of maximizing the test coverage by exercising the function under test (FUT) with different input parameters. Moreover, if a FUT interacts with the DOM of a web application, a \emph{test fixture} is expected as an additional input.

Generally, the problem of test data generation has been under an active study for the past three decades. In particular for JS, researchers have proposed several solutions based on: concolic execution (\emph{Jalangi}~\cite{sen2013jalangi}, \emph{Confix}~\cite{amin:ase15}), random generation (\emph{JSContest}~\cite{heidegger2010contract}), and the analysis of the whole web application (\emph{Artemis}~\cite{artemis2011}, Crawljax~\cite{mesbah2012crawling}). Out of those tools, only \emph{Confix} tries to construct a sufficient DOM fixture by translating native browser API calls into logical constraints. However, in the empirical evaluation~\cite{amin:ase15}, Confix only reached about 50\% in \emph{branch coverage} on the selected subjects. At the same time, search-based techniques have shown prominent results in achieving high branch coverage for imperative~\cite{wegener2001evolutionary}, object-oriented~\cite{fraser2011evosuite} and dynamic languages~\cite{irawan2016test, wibowo2015unit}, as well as for web~\cite{alshahwan2011automated} and mobile~\cite{mao2016sapienz} applications. But to the best of our knowledge, the problem of test data generation has never been tackled for JS with the help of the search-based approach. In this paper we are breaching this gap by introducing a \textbf{J}avascript \textbf{E}volution-based testing framework with \textbf{D}OM as an \textbf{I}nput, called \emph{JEDI}. We can summarize our contributions as follows: 
\begin{enumerate}[leftmargin=5mm]
\item The \emph{JEDI} framework is a novel JS unit testing tool using search-based techniques for generating test data. Notably, our testing framework is able to generate an arbitrary DOM input (both tags and attributes), which is syntactically valid and compliant with the recent HTML5 specification.   
\item Our test generation algorithm, called ``genetic with restart'', is trying to escape search plateau by restarting GA with a new target. Because the full-fledged JS data flow analysis~\cite{jang2009points} is hard to reach, the new target is chosen purely based on the preceding control flow behaviour in contrast with more popular techniques~\cite{ferguson1996chaining}.  
\item Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conducted an empirical validation of our framework and performed a significance study of the results. The validation has shown the effectiveness of the framework in covering target branches with a reasonable efficiency. Moreover, we showed that on the selected subjects \emph{JEDI} outperforms \emph{Confix} ....  
\end{enumerate}

The paper is structured as follows: In Section~\ref{sec.example}, we present a motivating example illustrating the challenges of generating test data for JS. The architecture of the \emph{JEDI} testing framework is explained in Section~\ref{sec.framework}. Section~\ref{sec.evaluation} presents empirical validation and indicates the validity threats. Related and future work are discussed in Section~\ref{sec:related.work}. Section~\ref{sec:concl} concludes the paper.

%% The state of JS 2017 survey~\footnote{\url{https://stateofjs.com/}} contains an extensive list of testing frameworks including Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma and QUnit. These frameworks support testing at various levels: unit, integration and functional. The survey also indicated a relatively low satisfaction rate of the current state of JS testing tools, scoring only 3.2 out of 5. As result, new testing frameworks such as Cypress and Storybook are constantly appearing to improve the situation. But despite this effort, test data generation continues to be poorly addressed by any of the widely used testing frameworks. The research community has actively been working towards making progress in this direction introducing such tools as Artemis~\cite{artemis2011}, JSConTest~\cite{heidegger2010contract}, Crawljax~\cite{mesbah2012crawling}, JSeft~\cite{mirshokraie2015jseft} and Jalangi~\cite{sen2013jalangi}. A recent study of client-side JS bugs~\cite{ocariza2017study} has shown that the majority (68\%) of web application faults are DOM-related and introduced in the JS code, and not in other components such as HTML or CSS. This fact indicates that yet a better testing approach is needed when it comes to unit testing of JS.

%% --------------------------------------------------------------------
\section{Motivating Example}
\label{sec.example}
%% --------------------------------------------------------------------

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
/*t dom */
function isGameFinished() {
  var obj = document.getElementById('sudoku'); |c \label{isGameFinished.getSudoku} |c
  var subDivs = obj.getElementsByTagName('DIV'); |c \label{isGameFinished.getDivs} |c
  var allOk = true;
  for (var no = 0; no < subDivs.length; no++) { |c \label{isGameFinished.inFor.begin} |c
    if (subDivs[no].className.indexOf('square') >= 0 |c \label{isGameFinished.if1.begin} |c 
        && !subDivs[no].style.backgroundColor) { 
      var spans = subDivs[no].getElementsByTagName('SPAN');
      if (spans[0].innerHTML != spans[1].innerHTML) { |c \label{isGameFinished.if2.begin} |c
        allOk = false; //target |c \label{isGameFinished.unfinished} |c
        break;}}}} |c \label{isGameFinished.outFor} |c
\end{lstlisting}
  \caption{JS function \texttt{isGameFinished.js} from sudoku}
  \label{code.isGameFinished}
\end{figure}

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs, language=HTML5]
<!-- T1: (5,5) -->            <!-- T2: (5,6), (7,15) -->
<html>                        <html>
 <body>                        <body>
  <div id='sudoku'>             <div id='sudoku'>
                                 <div></div>                           
  </div>                        </div> 
 </body>                       </body>    
</html>                       </html>

<!-- T3: (7,8), (10,14) -->   <!-- T4: (7,8), (10,11) -->
<html>                        <html>
 <body>                        <body>
  <div id='sudoku'>             <div id='sudoku'>
   <div class='square'>          <div class='square'>
    <span></span>                 <span></span> 
    <span></span>                 <span>TEST</span>
   </div>                        </div> 
  </div>                        </div>
 </body>                       </body>
</html>                       </html>
  \end{lstlisting}
  \caption{Input DOM arguments for \texttt{isGameFinished.js}}
  \label{fig.isGameFinished.tests}
\end{figure}

Let us consider the JS function \texttt{isGameFinished} in Figure~\ref{code.isGameFinished} taken from the web game \emph{sudoku}~\cite{sudoku}. This function validates a sudoku solution given by a user. First, it finds a DOM element representing the game field (line~\ref{isGameFinished.getSudoku}). Then, it collects all the child \emph{divs} (line~\ref{isGameFinished.getDivs}). Each \emph{div} element corresponds to a sudoku input square (line~\ref{isGameFinished.if1.begin}), and consists of two \emph{spans}. The first \emph{span} contains a user input, whereas the second one (hidden by default) stores an expected value for that sudoku square. Inside of a for-loop in lines~\ref{isGameFinished.inFor.begin}-\ref{isGameFinished.outFor}, we iterate over the whole collection of input squares. Once, we find a square with two unequal values (line~\ref{isGameFinished.if2.begin}), the loop is terminated and the game is considered unfinished (line~\ref{isGameFinished.unfinished}) and finished otherwise.

Suppose now that we would like to unit test this function in isolation to the rest of the application. We aim to maximize the branch coverage as a common testing criteria~\cite{zhu1997software}. That is, for each individual branch of the function under test (FUT) \texttt{isGameFinished}, we need to construct a test input that covers the branch and leads to the normal termination of the FUT. A JS function can take \emph{explicit} input arguments (in our case none), but it can also accept \emph{implicit} arguments such as a DOM state. In general, in order to produce a complete test for our function we have to construct both an appropriate DOM and input arguments.

%Our program has the following branches: $(6,15)$, $(6,7)$, $(7,9)$, $(7,14)$, $(10,11)$, $(10,13)$, $(16,17)$, $(16,18)$. Four tests in

Figure~\ref{fig.isGameFinished.tests} presents four tests that together provide the full coverage of the FUT. Let us take a look at how to construct the \texttt{T4} test, which covers the branch $(10,11)$, by following the structure of the FUT. Line~\ref{isGameFinished.getSudoku} expects the input DOM to have an element with an id \textquotesingle\texttt{sudoku}\textquotesingle. In order to enter the for-loop in line~\ref{isGameFinished.inFor.begin}, that element should have at least one child \emph{div} tag. The first if-condition in line~\ref{isGameFinished.if1.begin} requires this \emph{div} to be of the class \textquotesingle\texttt{square}\textquotesingle\ and also have no background color. The second if-condition (line~\ref{isGameFinished.if2.begin}) expects that the \emph{div} consists of two \emph{span} elements whose \emph{innerHtml} values are not equal. The resulting test \texttt{T4} in Figure~\ref{fig.isGameFinished.tests} meets all the above conditions.

As we have just observed, even for such a relatively simple example, the test data generation can be far from trivial procedure because it requires the deep understanding of the program semantics. In practice, web developers often have to deal with a large number of the JS frameworks and libraries freely available on the GitHub whose code is commonly untested and poorly documented. At the same time, unit testing is the most fundamental level of testing and the easiest to implement. It saves the developers from unexpected regressions and often documents the desired behavior of a program.

@@

\begin{figure}[t!]
  \centering 
\tikzset{
  % add this style to all tikzpicture environments
  every picture/.append style={
    % enable scaling of nodes
    transform shape,
    % set scale factor
    scale=1
  }
}
  \begin{sequencediagram}[font=\scriptsize]

@@

      \begin{sdblock}{\textbf{Genetic Loop}}{\scriptsize \hspace{8mm}$\text{branch} \in \text{branches}'$}
        \begin{sdblock}{\textbf{Random Loop}}{\scriptsize errors > 0}
          \begin{call}{server}{pop = Random(cinfo, config)}{validator}{errors}
          \end{call}
          \prelevel
        \end{sdblock}

@@

          \begin{sdblock}{\textbf{Crossover Loop}}{\scriptsize errors > 0}
            \begin{call}{server}{cross = Crossover(cinfo, pop)}{validator}{errors}
            \end{call}
            \prelevel
          \end{sdblock}

          \begin{sdblock}{\textbf{Mutation Loop}}{\scriptsize errors > 0}
            \begin{call}{server}{mut = Mutation(cinfo, pop)}{validator}{errors}
            \end{call}
            \prelevel
          \end{sdblock}

@@

\begin{algorithm}[t!]
  \caption{Initialization Phase}
  \label{alg.init}
  \algsetup{linenosize=\tiny}
  \scriptsize
  \DontPrintSemicolon
  \SetAlgoVlined
  \SetKwInOut{Input}{Input}
  \SetKwInOut{Output}{Output}
  \SetKwInOut{Require}{Require}
  \SetKwProg{Fn}{Function}{}{}
  \Input{JS file $fun.js$ with FUT and type annotation }
  \Output{Tuple $(cfg, branches, cinfo, ifun, sig)$}
  \Fn{$Initialize(fun.js)$}{

@@

\begin{figure}[t!]
\setlength{\grammarparsep}{3pt}
\footnotesize
\begin{grammar}
<JS Type Annotation> ::= /*t <Signature> */

@@

\subsubsection{Instrumentation}
\label{sub.sec.instrument}

Figure~\ref{code.isGameFinished.instr} presents the \texttt{isGameFinished} function after instrumentation. The global variable \texttt{trace} in line~\ref{isGameFinished.instr.trace} records the sequence of executed statements of the FUT. We have to trace all the statements because each one, in theory, can rise an exception. The \texttt{loopMap} variable in line~\ref{isGameFinished.instr.loopMap} captures the upper bound for the number for-loop iterations. If the exact bound is unknown, it is set to one. The \texttt{branchDistance} variable, computed for the \emph{problem node}, contains the distance~\cite{tracey1998automated} from the target branch.

%TODO: maybe expand this section with the JS type casting for branch distance computation

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
function isGameFinished() {
 trace.push(1);var obj = document.getElementById("sudoku"); |c \label{isGameFinished.instr.trace} |c
 trace.push(2);var subDivs = obj.getElementsByTagName("DIV");
 var allOk = true;
 loopMap[3] = function () { |c \label{isGameFinished.instr.loopMap} |c
   var no = 0;
   if (subDivs.length || subDivs.length == 0) {
     return Math.abs(no - subDivs.length);
   }; return 1}();
 trace.push(3);
 for (var no = 0; no < subDivs.length; no++) {
  trace.push(4);
  if (subDivs[no].className.indexOf("square") >= 0 &&
      !subDivs[no].style.backgroundColor) {
   trace.push(5); branchDistance.push({ label: 3, 
    distance: Math.min(
     abs(subDivs[no].className.indexOf("square"),0)+_K_ 
     absZero(subDivs[no].style.backgroundColor))
   });
   trace.push(6);var spans=subDivs[no].getElementsByTagName("SPAN");
   trace.push(7);
   if (spans[0].innerHTML != spans[1].innerHTML) {
    trace.push(8); branchDistance.push({ label: 7, 
     distance: abs(spans[0].innerHTML, spans[1].innerHTML)
    });
    allOk = false; break;
   } else {
   trace.push(9);branchDistance.push({label:7,distance:_K_});
   }
  } else {
   trace.push(10); branchDistance.push({ label: 3,
    distance: 
     abs(subDivs[no].className.indexOf("square"), 0) + 
     absNegZero(subDivs[no].style.backgroundColor)
   });}}
 trace.push(-1);
}
\end{lstlisting}
  \caption{Instrumented version of \texttt{isGameFinished.js} }
  \label{code.isGameFinished.instr}
\end{figure}

@@

In some cases, when we do not get enough guidance from a fitness function, the genetic algorithm converges to a local minimum~\cite{mcminn2004search}. For example, if one branch is dominated by another, the coverage of the former depends on the coverage of the latter. Unless this relation is encoded in the fitness function somehow, we can not learn about it accidentally. In such a situation, we find ourselves being trapped on a flat space, with a small chance to escape. In order to overcome this problem, we propose a novel --- to the best of our knowledge --- solution based on the idea of \emph{archive convergence}. If during a certain history window, the fitness values of the respective archives have converged to a fixed point, we deem the whole history converged and restart the genetic search with a new search target. At the end of the evolution loop (line~\ref{alg.gen.has.converged}), we verify the convergence of the history within the window of $size_H$. If that condition is true (line~\ref{alg.gen.loop.end}), the genetic algorithm is restarted with a new search target. It consists of the original branch $b_n$ prefixed with a new branch $b$ from which $b_n$ depends on (line~\ref{alg.gen.new.target}). Every time the history converges, we produce a new search target by eventually exhausting all possible executions leading to $b_n$. In practice, we only consider extensions upto length two.

@@

\begin{table}
  \caption{Genetic Operation for the JS Types}
  \label{tbl.gen.oper.js.types}
    \footnotesize
  \begin{tabular}{p{1cm}|p{5mm}|p{6cm}}
    \toprule
    \textbf{Operation} & \textbf{Type} & \textbf{Description} \\
    \hline
    Random    & Bool   & choose at random \{\texttt{true}, \texttt{false}\} \\
    Random    & Int    & choose at random [-10, 10] or constants with the ratio 4:1\\
    Random    & Float  & choose at random [-10.0, 10.0] or constants with the ratio 4:1\\
    Random    & String & >5 chars from $\texttt{[a-zA-Z0-9]}$ or constants with the ratio 4:1\\
    Random    & Array  & array of a respective type upto length five\\
    \hline
    Crossover & Bool   & pick at random one out of two given booleans \\
    Crossover & Int    & pick at random one out of two given integers\\
    Crossover & Float  & pick at random one out of two given floats\\
    Crossover & String & pick at random one out of two single point crossovers\\
    Crossover & Array  & similar crossover operator as for strings\\
    \hline
    Mutation  & Bool   & flip the boolean constant\\
    Mutation  & Int    & random integer, $\pm 1$ or $\pm 10$\\
    Mutation  & Float  & random float, $\pm 0.1$, $\pm 1$ or $\pm 10$\\
    Mutation  & String & random string, remove char or apply pred(suc) char\\
    Mutation  & Array  & Mutate a random element in array according to its type\\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{figure}[t]
  \begin{haskell}
instance Arbitrary Html where
  arbitrary = evalStateT generateHTML defaultState

>>>>

@@

\begin{figure}[!t]
  \centering
\tikzset{
  % add this style to all tikzpicture environments
  every picture/.append style={
    % enable scaling of nodes
    transform shape,
    % set scale factor
    scale=1
  }
}
  \begin{sequencediagram}[font=\scriptsize]

@@

      \begin{sdblock}{\textbf{Genetic Loop}}{\scriptsize \hspace{8mm}$\text{branch} \in \text{branches}'$}
        \begin{sdblock}{\textbf{Random Loop}}{\scriptsize errors $>$ 0}
          \begin{call}{server}{pop = Random(cinfo, config)}{validator}{errors}
          \end{call}
          \prelevel
        \end{sdblock}

@@

          \begin{sdblock}{\textbf{Crossover Loop}}{\scriptsize errors $>$ 0}
            \begin{call}{server}{cross = Crossover(cinfo, pop)}{validator}{errors}
            \end{call}
            \prelevel
          \end{sdblock}

          \begin{sdblock}{\textbf{Mutation Loop}}{\scriptsize errors $>$ 0}
            \begin{call}{server}{mut = Mutation(cinfo, pop)}{validator}{errors}
            \end{call}
            \prelevel
          \end{sdblock}

@@

\begin{algorithm}[!t]
  \caption{Initialization Phase}
  \label{alg.init}
  \algsetup{linenosize=\tiny}
  \scriptsize
  \DontPrintSemicolon
  \SetAlgoVlined
  \SetKwInOut{Input}{Input}
  \SetKwInOut{Output}{Output}
  \SetKwInOut{Require}{Require}
  \SetKwProg{Fn}{Function}{}{}
  \Input{JS file $fun.js$ with FUT and type annotation }
  \Output{Tuple $(cfg, branches, cinfo, ifun, sig)$}
  \Fn{$Initialize(fun.js)$}{

@@

\begin{figure}[!t]
\setlength{\grammarparsep}{3pt}
\scriptsize
\begin{grammar}
<JS Type Annotation> ::= /*t <Signature> */

@@

%% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
%%        LEFT OUT DUE TO LACK OF SPACE
%% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
%% \subsubsection{Instrumentation}
%% \label{sub.sec.instrument}

%% Figure~\ref{code.isGameFinished.instr} presents the \texttt{isGameFinished} function after instrumentation. The global variable \texttt{trace} in line~\ref{isGameFinished.instr.trace} records the sequence of executed statements of the FUT. We have to trace all the statements because each one, in theory, can rise an exception. The \texttt{loopMap} variable in line~\ref{isGameFinished.instr.loopMap} captures the upper bound for the number for-loop iterations. If the exact bound is unknown, it is set to one. The \texttt{branchDistance} variable, computed for the \emph{problem node}, contains the distance~\cite{tracey1998automated} from the target branch.

%% %TODO: maybe expand this section with the JS type casting for branch distance computation

%% \begin{figure}[!t]
%%   \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
%% function isGameFinished() {
%%  trace.push(1);var obj = document.getElementById("sudoku"); |c \label{isGameFinished.instr.trace} |c
%%  trace.push(2);var subDivs = obj.getElementsByTagName("DIV");
%%  var allOk = true;
%%  loopMap[3] = function () { |c \label{isGameFinished.instr.loopMap} |c
%%    var no = 0;
%%    if (subDivs.length || subDivs.length == 0) {
%%      return Math.abs(no - subDivs.length);
%%    }; return 1}();
%%  trace.push(3);
%%  for (var no = 0; no < subDivs.length; no++) {
%%   trace.push(4);
%%   if (subDivs[no].className.indexOf("square") >= 0 &&
%%       !subDivs[no].style.backgroundColor) {
%%    trace.push(5); branchDistance.push({ label: 3,
%%     distance: Math.min(
%%      abs(subDivs[no].className.indexOf("square"),0)+_K_
%%      absZero(subDivs[no].style.backgroundColor))
%%    });
%%    trace.push(6);var spans=subDivs[no].getElementsByTagName("SPAN");
%%    trace.push(7);
%%    if (spans[0].innerHTML != spans[1].innerHTML) {
%%     trace.push(8); branchDistance.push({ label: 7,
%%      distance: abs(spans[0].innerHTML, spans[1].innerHTML)
%%     });
%%     allOk = false; break;
%%    } else {
%%    trace.push(9);branchDistance.push({label:7,distance:_K_});
%%    }
%%   } else {
%%    trace.push(10); branchDistance.push({ label: 3,
%%     distance:
%%      abs(subDivs[no].className.indexOf("square"), 0) +
%%      absNegZero(subDivs[no].style.backgroundColor)
%%    });}}
%%  trace.push(-1);
%% }
%% \end{lstlisting}
%%   \caption{Instrumented version of \texttt{isGameFinished.js} }
%%   \label{code.isGameFinished.instr}
%% \end{figure}

@@

However, ff during a certain history window, the fitness values of the respective archives have converged to a fixed point, we deem the whole history converged and restart the genetic search with a new search target. At the end of the evolution loop (line~\ref{alg.gen.has.converged}), we verify the convergence of the history within the window of $size_H$. If that condition is true (line~\ref{alg.gen.loop.end}), the genetic algorithm is restarted with a new search target. It consists of the original branch $b_n$ prefixed with a new branch $b$ from which $b_n$ depends on (line~\ref{alg.gen.new.target}). Every time the history converges, we produce a new search target by eventually exhausting all possible executions leading to $b_n$. In practice, we only consider extensions upto length two.

In some cases, we get insufficient guidance from a fitness function because the FUT contains \emph{flag variables}~\cite{baresel2004evolutionary}, nested predicates~\cite{mcminn2005testability}, or unstructured control flow~\cite{hierons2005branch}. Researches have devised several techniques, such as \emph{chaining approach}~\cite{ferguson1996chaining,mcminn2006evolutionary} and \emph{testability transformation}~\cite{korel2005data}, to overcome with these problems. The common idea behind of these techniques is to incorporate data dependency into the search. In our case, such an analysis for JS is hard to implement~\cite{jang2009points}. So we decided to build our search algorithm purely on the knowledge about control-flow dependencies. The downside of this idea is the lack of precision, i.e some of the node sequences are fruitless to tackle. Nevertheless, it was sufficient to find solutions for our experimental subjects.

@@

\begin{table}[!t]
  \caption{Genetic Operation for the JS Types}
  \label{tbl.gen.oper.js.types}
  \scriptsize
  \centering
  \begin{tabular}{p{1cm}|p{5mm}|p{6cm}}
    \toprule
    \textbf{Operation} & \textbf{Type} & \textbf{Description} \\
    \hline
    Random    & Bool   & choose at random \{\texttt{true}, \texttt{false}\} \\
    Random    & Int    & choose (4:1) at random [-10, 10] or constants\\
    Random    & Float  & choose (4:1) at random [-10.0, 10.0] or constants\\
    Random    & String & $>5$ chars from $\texttt{[a-zA-Z0-9]}$ or constants (4:1)\\
    Random    & Array  & array of a respective type upto length five\\
    \hline
    Crossover & Bool   & pick at random one out of two given booleans \\
    Crossover & Int    & pick at random one out of two given integers\\
    Crossover & Float  & pick at random one out of two given floats\\
    Crossover & String & pick at random one out of two single point crossovers\\
    Crossover & Array  & similar crossover operator as for strings\\
    \hline
    Mutation  & Bool   & flip the boolean constant\\
    Mutation  & Int    & random integer, $\pm 1$ or $\pm 10$\\
    Mutation  & Float  & random float, $\pm 0.1$, $\pm 1$ or $\pm 10$\\
    Mutation  & String & random string, remove char or apply pred(suc) char\\
    Mutation  & Array  & Mutate a random element in array according to its type\\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{figure}[!t]
  \centering
  \begin{haskell}
instance Arbitrary Html where
  arbitrary = evalStateT generateHTML defaultState
